---
title: "P8105 Homework 3"
author: "Jesse R. Ames"
date: "10/20/2021"
output: github_document
---

```{r load}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

## Problem 1

(a) How many aisles are there, and which aisles are the most items ordered from?

```{r 1a}
#How many isles are there?
instacart %>%
  pull("aisle") %>%
  unique() %>%
  length()

#Which aisles are the most items ordered from?
instacart %>%
  pull("aisle") %>%
  janitor::tabyl() %>%
  arrange(desc(n)) %>%
  head(10) #Top 10 aisles by number of items ordered
```

(b) Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r 1b}
instacart %>%
  pull("aisle") %>%
  janitor::tabyl() %>%
  filter(n > 10000) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = reorder(.,n) , y = n)) +
  geom_col() +
  coord_flip() +
  labs(y = "Number of items ordered",
       x = "Aisle",
       title = "Aisles with more than 10,000 ordered items")
```

(c) Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r 1c}
instacart %>%
  filter(aisle %in% c(
    "baking ingredients",
    "dog food care",
    "packaged vegetables fruits"
  )) %>%
  select(aisle, product_name) %>%
  pivot_wider(names_from = aisle,
              values_from = product_name,
              values_fn = length) %>%
  janitor::clean_names()
  #select(baking_ingredients, dog_food_care, packaged_vegetables_fruits)
  # group_by(aisle, product_name) %>%
  # tally(sort = TRUE)
```

(d) Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

## Problem 2

(a) Clean the data.

```{r}
data("brfss_smart2010")
brfss <- brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  #pull(response) %>% unique() -- We have responses from "Excellent" to "Poor"
  mutate(
    response = factor(response,
       levels = c("Poor",
                  "Fair",
                  "Good",
                  "Very good",
                  "Excellent")
      )
    )
```

(b) In 2002, which states were represented at 7 or more locations?

```{r}
brfss %>%
  filter(year == 2002) %>%
  pull(locationdesc) %>% unique()
  # group_by(locationabbr) %>%
  # summarise(locationabbr,locationdesc) %>%
  # janitor::tabyl(locationabbr) %>% pull(n) %>% sum()
```

(c) Construct a dataset that is limited to "Excellent" responses and contains year, state, and a variable that averages the `data_value` across locations. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the `geom_line` geometry and `group` aesthetic will help).

```{r}
brfss %>%
  filter(response == "Excellent") %>%
  group_by(locationabbr, year) %>%
  summarise(mean_data_value = mean(data_value)) %>%
  ggplot(aes(x = year, y = mean_data_value, group = locationabbr)) +
  geom_line() #Not the most useful plot
```

(d) Make a two-panel plot showing, for the years 2006, and 2010, distribution of `data_value` for responses (“Poor” to “Excellent”) among locations in NY State

```{r}
brfss %>%
  filter(locationabbr == "NY", year %in% c(2006,2010)) %>%
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(.~year)
```

